---
title: "CellChat_BC_Project"
author: "Beth Cragg"
date: "2025-06-23"
output:
  word_document: default
  html_document: default
  pdf_document: default
#CellChat Analysis - BC Research Project v2.0
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(CellChat)
library(Seurat)
library(cowplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(CellChat)
#source("Utils.R")
library(tidyverse)
library(NMF)
library(circlize)
library(ComplexHeatmap)
```

# Run preprocessing for each timepoint separately
```{r, echo=false}

#Read in final data files
control_seurat <- readRDS(file = "2025_MSc_control.rds")
h24_seurat <- readRDS(file = "2025_MSc_24.rds")
h48_seurat <- readRDS(file = "2025_MSc_48.rds")

#  - - - - - CONTROL - - - - - - 
# Extract normalized data & metadata
data.input.control <- GetAssayData(control_seurat, assay = "RNA", layer= "data")  # log-normalized counts
meta.control <- data.frame(labels = control_seurat$cell_type_broad, row.names = colnames(control_seurat))
rownames(meta.control) <- colnames(data.input.control)

cellchat.control <- createCellChat(object = data.input.control, meta = meta.control, group.by = "labels")
cellchat.control <- addMeta(cellchat.control, meta = meta.control)
cellchat.control <- setIdent(cellchat.control, ident.use = "labels")

# Assign CellChatDB (human or mouse depending on your data)
cellchat.control@DB <- CellChatDB.mouse

# Preprocess
cellchat.control <- subsetData(cellchat.control)
cellchat.control <- identifyOverExpressedGenes(cellchat.control, do.fast = FALSE)
cellchat.control <- identifyOverExpressedInteractions(cellchat.control)
cellchat.control <- computeCommunProb(cellchat.control)
cellchat.control <- filterCommunication(cellchat.control, min.cells=10)
cellchat.control <- computeCommunProbPathway(cellchat.control)
cellchat.control <- aggregateNet(cellchat.control)
cellchat.control <- netAnalysis_computeCentrality(cellchat.control)



#  - - - - - 24 HRS - - - - - - 
# Extract normalized data & metadata
data.input.h24 <- GetAssayData(h24_seurat, assay = "RNA", layer= "data")  # log-normalized counts
meta.h24 <- data.frame(labels = h24_seurat$cell_type_broad)
rownames(meta.h24) <- colnames(data.input.h24)

cellchat.h24 <- createCellChat(object = data.input.h24, meta = meta.h24, group.by = "labels")
cellchat.h24 <- addMeta(cellchat.h24, meta = meta.h24)
cellchat.h24 <- setIdent(cellchat.h24, ident.use = "labels")

# Assign CellChatDB (human or mouse depending on your data)
cellchat.h24@DB <- CellChatDB.mouse

# Preprocess
cellchat.h24 <- subsetData(cellchat.h24)
cellchat.h24 <- identifyOverExpressedGenes(cellchat.h24, do.fast = FALSE)
cellchat.h24 <- identifyOverExpressedInteractions(cellchat.h24)
cellchat.h24 <- computeCommunProb(cellchat.h24)
cellchat.h24 <- filterCommunication(cellchat.h24, min.cells=10)
cellchat.h24 <- computeCommunProbPathway(cellchat.h24)
cellchat.h24 <- aggregateNet(cellchat.h24)
cellchat.h24 <- netAnalysis_computeCentrality(cellchat.h24)



#  - - - - - 48 HRS - - - - - - 
# Extract normalized data & metadata
data.input.h48 <- GetAssayData(h48_seurat, assay = "RNA", layer= "data")  # log-normalized counts
meta.h48 <- data.frame(labels = h48_seurat$cell_type_broad)
rownames(meta.h48) <- colnames(data.input.h48)

cellchat.h48 <- createCellChat(object = data.input.h48, meta = meta.h48, group.by = "labels")
cellchat.h48 <- addMeta(cellchat.h48, meta = meta.h48)
cellchat.h48 <- setIdent(cellchat.h48, ident.use = "labels")

# Assign CellChatDB (human or mouse depending on your data)
cellchat.h48@DB <- CellChatDB.mouse

# Preprocess
cellchat.h48 <- subsetData(cellchat.h48)
cellchat.h48 <- identifyOverExpressedGenes(cellchat.h48, do.fast = FALSE)
cellchat.h48 <- identifyOverExpressedInteractions(cellchat.h48)
cellchat.h48 <- computeCommunProb(cellchat.h48)
cellchat.h48 <- filterCommunication(cellchat.h48, min.cells=10)
cellchat.h48 <- computeCommunProbPathway(cellchat.h48)
cellchat.h48 <- aggregateNet(cellchat.h48)
cellchat.h48 <- netAnalysis_computeCentrality(cellchat.h48)

object.list <- list(
  Cycling = cellchat.control,
  H24 = cellchat.h24,
  H48 = cellchat.h48
)

cellchat <- mergeCellChat(object.list, add.names = names(object.list))

```


## Compare the total number of interactions and interaction strength
To answer on question on whether the cell-cell communication is enhanced or not, CellChat compares the the total number of interactions and interaction strength of the inferred cell-cell communication networks from different biological conditions.

##Also generate scatter plots to map senders/receivers for each timepoint
```{r, echo=false}
#analyse immune and stromal separately
immune <- subsetCellChat(cellchat, group.by = "labels", idents.use = c("Neut","Mono/Macs","NK","DC","Tc","Bc"))
stromal <- subsetCellChat(cellchat, group.by = "labels", idents.use = c("Fibroblasts", "MET"))

# Compare interactions in immune cells
gg_immune1 <- compareInteractions(immune, show.legend = FALSE, group = c(1:3), title="Number of interactions (immune cells only)")
gg_immune2 <- compareInteractions(immune, show.legend = FALSE, group = c(1:3), measure = "weight", title="Interaction strength (immune cells only)")

# Compare interactions in all cells
gg1 <- compareInteractions(cellchat, show.legend = FALSE, group = c(1:3), title="Number of interactions")
gg2 <- compareInteractions(cellchat, show.legend = FALSE, group = c(1:3), measure = "weight", title="Interaction strength")
gg1+gg2

# Compare interactions in stromal cells
gg_stromal1 <- compareInteractions(stromal, show.legend = FALSE, group = c(1:3), title="Number of interactions (stromal cells only)")
gg_stromal2 <- compareInteractions(stromal, show.legend = FALSE, group = c(1:3), measure = "weight", , title="Interaction strength (stromal cells only)")

# Plot side by side
gg_immune1+gg_immune2
gg_stromal1+gg_stromal2

#Scatter plots
scatter.control <- netAnalysis_signalingRole_scatter(cellchat.control, title="Signalling Roles by Cell Type - Cycling")
scatter.24 <- netAnalysis_signalingRole_scatter(cellchat.h24, title="Signalling Roles by Cell Type - 24hr")
scatter.48 <- netAnalysis_signalingRole_scatter(cellchat.h48, title="Signalling Roles by Cell Type - 48hr")

get_plot_data <- function(p) ggplot_build(p)$data[[1]]
all_data <- do.call(rbind, lapply(list(scatter.control, scatter.24, scatter.48), get_plot_data))
xr <- range(all_data$x)
yr <- range(all_data$y)

scatter.control + coord_cartesian(xlim=xr, ylim=yr)
scatter.24      + coord_cartesian(xlim=xr, ylim=yr)
scatter.48      + coord_cartesian(xlim=xr, ylim=yr)

```


#Compare interaction strength by cell type
#red = increases over time
#blue = decreases over time
```{r, echo=false}
netVisual_diffInteraction(cellchat, comparison = c(1, 2), weight.scale = T, measure = "weight", title.name = "Communication Changes: Cycling vs 24hr")
title(main = "Communication Changes: Cycling vs 24hr", cex.main = 1.2)
netVisual_diffInteraction(cellchat, comparison = c(1, 3), weight.scale = T, measure = "weight")
netVisual_diffInteraction(cellchat, comparison = c(2, 3), weight.scale = T, measure = "weight")
```



#Circle plots for mapping interactions links
#Heatmaps for analysing interaction strength between cell types at the different timepoints

```{r, echo=false}

# Cycling timepoint
weight_mat <- cellchat.control@net$weight
weight_mat[weight_mat <= 0.5] <- 0  #filter high confidence interactions only

# Now plot only high-confidence edges
netVisual_circle(
  weight_mat,
  vertex.weight = as.numeric(table(cellchat.control@idents)),
  weight.scale = TRUE,
  label.edge = FALSE
)
title(main = "Cycling (Interaction Strength >0.5)", cex.main = 1.2)


# 24h timepoint
weight_mat <- cellchat.h24@net$weight
weight_mat[weight_mat <= 0.5] <- 0  #filter high confidence interactions only

# Now plot only high-confidence edges
netVisual_circle(
  weight_mat,
  vertex.weight = as.numeric(table(cellchat.h24@idents)),
  weight.scale = TRUE,
  label.edge = FALSE
)
title(main = "24hr (Interaction Strength >0.5)", cex.main = 1.2)

# 48h timepoint
weight_mat <- cellchat.h48@net$weight
weight_mat[weight_mat <= 0.5] <- 0  #filter high confidence interactions only

# Now plot only high-confidence edges
netVisual_circle(
  weight_mat,
  vertex.weight = as.numeric(table(cellchat.h48@idents)),
  weight.scale = TRUE,
  label.edge = FALSE
)
title(main = "48hr (Interaction Strength >0.5)", cex.main = 1.2)

#Heatmaps
netVisual_heatmap(cellchat.control, title.name = "Number of interactions - Cycling")
netVisual_heatmap(cellchat.h24, title.name = "Number of interactions - 24hr")
netVisual_heatmap(cellchat.h48, title.name = "Number of interactions - 48hr")

```

#What are the top pathways by timepoint?

```{r}
top_cycling <- head(cellchat.control@netP$pathways, 20)
top_24 <- head(cellchat.h24@netP$pathways, 20)
top_48 <- head(cellchat.h48@netP$pathways, 20)
all_pathways <- unique(c(top_cycling, top_24, top_48))
pathway_counts <- table(c(top_cycling, top_24, top_48))
top_20_merged <- names(sort(pathway_counts, decreasing = TRUE)[1:20])

# OUTGOING ROLE across all timepoints using same top pathways
print(
  netAnalysis_signalingRole_heatmap(
    cellchat.control,
    pattern = "outgoing",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "Cycling (Top 20 merged pathways)"
  )
)

print(
  netAnalysis_signalingRole_heatmap(
    cellchat.h24,
    pattern = "outgoing",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "24h (Top 20 merged pathways)"
  )
)

print(
  netAnalysis_signalingRole_heatmap(
    cellchat.h48,
    pattern = "outgoing",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "48h (Top 20 merged pathways)"
  )
)

# ✅ INCOMING ROLE across all timepoints
print(
  netAnalysis_signalingRole_heatmap(
    cellchat.control,
    pattern = "incoming",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "Cycling (Top 20 merged pathways)"
  )
)

print(
  netAnalysis_signalingRole_heatmap(
    cellchat.h24,
    pattern = "incoming",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "24h (Top 20 merged pathways)"
  )
)

print(
  netAnalysis_signalingRole_heatmap(
    cellchat.h48,
    pattern = "incoming",
    signaling = top_20_merged,
    width = 6, height = 6,
    title = "48h (Top 20 merged pathways)"
  )
)

```

#Identify clusters of pathways based off of functional similarity

```{r, echo=false}
library(reticulate)
reticulate::py_install("umap-learn", envname = NULL, pip = TRUE)
py_module_available("umap")
cellchat <- computeNetSimilarityPairwise(cellchat, type = "functional")
cellchat <- netEmbedding(cellchat, type = "functional")
cellchat <- netClustering(cellchat, type = "functional")

# Visualize in 2D embedding
netVisual_embeddingPairwise(cellchat, type = "functional", label.size = 3.5)
netVisual_embeddingPairwiseZoomIn(cellchat, type = "functional", nCol = 2)
rankSimilarity(cellchat, type = "functional")

# Extract cluster assignment
cluster_assignments <- cellchat@netP$similarity$functional$group

# Create dataframe of cluster allocation
cluster_assignments_vec <- unlist(cluster_assignments)
df_clusters <- data.frame(
  pathway_timepoint = names(cluster_assignments_vec),
  cluster = cluster_assignments_vec
)

# Split into timepoint + pathway
df_clusters <- df_clusters %>%
  separate(pathway_timepoint, into = c("pathway","timepoint"), sep = "--", extra = "merge")

# Remove the fixed "1-2-3." prefix from pathway names
df_clusters <- df_clusters %>%
  mutate(pathway = gsub("^1-2-3\\.", "", pathway))

df_clusters <- df_clusters %>%
  mutate(cluster = as.character(cluster)) %>%
  mutate(cluster = recode(cluster,
    "3" = "1",
    "1" = "2",
    "4" = "3",
    "2" = "4"
  ))

# Then convert to factor with new levels in desired order (1 to 4)
df_clusters$cluster <- factor(df_clusters$cluster, levels = c("1", "2", "3", "4"))

# If you have cluster <-> color mapping (from your scatter plots):
cluster_colors <- c("1" = "#FBB4AE", "2" = "#B3DE69", "3" = "#80B1D3", "4" = "#DDA0DD")

# Count num of pathways per cluster per timepoint
counts <- df_clusters %>%
  group_by(cluster, timepoint) %>%
  summarise(num_pathways = n(), .groups = "drop")

# 1. Count number of pathways per cluster
pathways_per_cluster <- df_clusters %>%
  group_by(cluster) %>%
  summarise(n_pathways = n(), .groups = "drop")

# 2. Count number of timepoints per cluster
timepoints_per_cluster <- df_clusters %>%
  group_by(cluster, timepoint) %>%
  summarise(n_timepoints = n_distinct(timepoint), .groups = "drop")

# 3. List of pathways per timepoint per cluster
pathways_per_timepoint_cluster <- df_clusters %>%
  group_by(cluster, timepoint) %>%
  summarise(pathways = paste(pathway, collapse = ", "), .groups = "drop")

# Plot number of pathways per cluster
ggplot(pathways_per_cluster, aes(x = factor(cluster), y = n_pathways, fill = factor(cluster))) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Pathways per Cluster", x = "Cluster", y = "Number of Pathways") +
  theme_minimal() +
  scale_fill_manual(values = cluster_colors) 

# Plot with cluster order/color reflecting CellChat results
ggplot(counts, aes(x = cluster, y = num_pathways, fill = cluster)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ timepoint) +
  scale_fill_manual(values = cluster_colors) +
  labs(
    x = "Cluster",
    y = "Number of Pathways",
    title = "Number of Pathways per Cluster per Timepoint"
  ) +
  theme_minimal()

# Step 1: Create the full pathway-timepoint name in df_clusters
df_clusters <- df_clusters %>%
  mutate(full_name = paste(pathway, timepoint, sep = "--"))

comm_probabilities <- cellchat@netP$similarity$functional$matrix[[1]]

# Calculate the average communication probability for each pathway (mean across the row)
comm_probabilities_avg <- apply(comm_probabilities, 1, mean)

# Create a data frame with the pathway names and corresponding communication probabilities
df_comm_prob <- data.frame(
  pathway = rownames(comm_probabilities),
  comm_probability = comm_probabilities_avg
)

# Split into timepoint + pathway
df_comm_prob <- df_comm_prob %>%
  separate(pathway, into = c("pathway","timepoint"), sep = "--", extra = "merge")

# Remove the fixed "1-2-3." prefix from pathway names
df_clusters <- df_clusters %>%
  mutate(pathway = gsub("^1-2-3\\.", "", pathway))

df_comm_prob <- data.frame(
  pathway_timepoint = rownames(df_comm_prob),
  df_comm_prob,  # Add the rest of the columns
  stringsAsFactors = FALSE
)

df_clusters <- df_clusters %>%
  left_join(df_comm_prob, by = c("full_name" = "pathway_timepoint"))

df_clusters <- df_clusters %>%
  select(-pathway.y, -timepoint.y)

#write.csv(df_clusters, file="df_functional_clusters_26.7.2025.csv")

df_clusters_sorted <- df_clusters %>%
  arrange(cluster, desc(comm_probability))  # Sort pathways by communication probability (descending)

# Step 2: Select top 3 pathways per cluster
top_3_pathways <- df_clusters_sorted %>%
  group_by(cluster) %>%
  slice_head(n = 3)  # Select top 3 rows per cluster

# Step 3: Check the top 3 pathways per cluster
print(top_3_pathways)

```

#View top 10 pathways by communication probability for each cluster

```{r}
# Sort and get top 10 pathways by communication probability within each cluster
top_10_cluster_pathways <- df_clusters %>%
  group_by(cluster) %>%
  arrange(cluster, desc(comm_probability)) %>%
  slice_head(n = 10)  # Get the top 10 pathways per cluster

# View the top 10 pathways per cluster
print(top_10_cluster_pathways)
max_comm_prob <- max(top_10_cluster_pathways$comm_probability)

# Sort and get top 10 pathways by communication probability within each timepoint
top_10_timepoint_pathways <- df_clusters %>%
  group_by(timepoint.x) %>%
  arrange(timepoint.x, desc(comm_probability)) %>%
  slice_head(n = 10)  # Get the top 10 pathways per timepoint

# View the top 10 pathways per timepoint
print(top_10_timepoint_pathways)


# Loop through each cluster and plot separately
for(cluster_id in unique(top_10_cluster_pathways$cluster)) {
  
  # Filter data for the current cluster
  cluster_data <- top_10_cluster_pathways %>%
    filter(cluster == cluster_id)
  
  # Create the plot for the current cluster
  p <- ggplot(cluster_data, aes(x = reorder(pathway.x, comm_probability), y = comm_probability, fill = factor(cluster))) +
    geom_bar(stat = "identity") +
    labs(title = paste("Top 10 Pathways - Cluster", cluster_id),
         x = "Pathway", y = "Communication Probability") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("1" = "#66c2a5", "2" = "#fc8d62", "3" = "#8da0cb", "4" = "#e78ac3")) +
    ylim(0, max_comm_prob)  + theme(legend.position="none")
  print(p)
}

# Loop through each cluster and plot separately
for(timepoint_id in unique(top_10_timepoint_pathways$timepoint.x)) {
  
  # Filter data for the current cluster
  timepoint_data <- top_10_timepoint_pathways %>%
    filter(timepoint.x == timepoint_id)
  
  # Create the plot for the current cluster
  # Plot top 10 pathways by communication probability for each timepoint (individual plots)
  p <- ggplot(timepoint_data, aes(x = reorder(pathway.x, comm_probability), y = comm_probability, fill = factor(timepoint.x))) +
    geom_bar(stat = "identity") +
    labs(title = "Top 10 Pathways by Communication Probability per Timepoint", x = "Pathway", y = "Communication Probability") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("Cycling" = "#66c2a5", "H24" = "#fc8d62", "H48" = "#8da0cb"))+
    ylim(0, max_comm_prob)
  print(p)
}

# Correlation between communication probability and other factors (e.g., cluster)
cor_analysis <- df_clusters %>%
  group_by(cluster) %>%
  summarise(correlation = cor(comm_probability, as.numeric(factor(cluster))))

print(cor_analysis)

# Boxplot showing distribution of communication probabilities by cluster
ggplot(df_clusters, aes(x = factor(cluster), y = comm_probability, fill = factor(cluster))) +
  geom_boxplot() +
  labs(title = "Distribution of Communication Probabilities by Cluster", x = "Cluster", y = "Communication Probability") +
  theme_minimal() +
  scale_fill_manual(values = c("1" = "#66c2a5", "2" = "#fc8d62", "3" = "#8da0cb", "4" = "#e78ac3"))


```


#Identify clusters of pathways based off of structural similarity
```{r, echo=false}
cellchat <- computeNetSimilarityPairwise(cellchat, type = "structural")
cellchat <- netEmbedding(cellchat, type = "structural")
cellchat <- netClustering(cellchat, type = "structural")
# Visualization in 2D-space
netVisual_embeddingPairwise(cellchat, type = "structural", label.size = 3.5)
netVisual_embeddingPairwiseZoomIn(cellchat, type = "structural", nCol = 2)

```

#Compare the overall information flow of each signaling pathway

```{r, echo=false, fig.width=10, fig.height=10}
gg1 <- rankNet(cellchat, mode = "comparison", stacked = T, do.stat = TRUE)
gg2 <- rankNet(cellchat, mode = "comparison", stacked = F, do.stat = TRUE)
gg1
gg2

netVisual_clustering(cellchat, type = "functional", cluster.id = 1) # for Cluster 1
netVisual_clustering(cellchat, type = "functional", cluster.id = 2) # for Cluster 2
netVisual_clustering(cellchat, type = "functional", cluster.id = 3) # for Cluster 3
netVisual_clustering(cellchat, type = "functional", cluster.id = 4) # for Cluster 4
```
#Compare outgoing (or incoming) signaling associated with each cell population

```{r, fig.align = "center"}
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)

ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 6, color.heatmap = "OrRd")
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 6, color.heatmap = "OrRd")
ht1
ht2
```

```{r, fig.width=12,fig.height = 6, fig.wide = TRUE, fig.align = "center"}
# Chord diagram
pathways.show <- c("MPZ") 
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
  netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "chord", signaling.name = paste(pathways.show, names(object.list)[i]))
}
```
```{r}
pathways.show <- c("MPZ", "NCAM")
weight.max <- sapply(pathways.show, function(pathway) {
  getMaxWeight(object.list, slot.name = "netP", attribute = pathway)
})

# Loop through each dataset (timepoint) and each pathway
for (i in 1:length(object.list)) {  # Loop over datasets
  for (j in 1:length(pathways.show)) {  # Loop over pathways
    # Generate pathway visualization for each dataset and pathway
    netVisual_aggregate(
      object.list[[i]], 
      signaling = pathways.show[j], 
      layout = "circle",  # You can also try "chord" for different layouts
      edge.weight.max = weight.max[j],  # Use the correct weight max for each pathway
      edge.width.max = 10,  # Max width of the edges
      signaling.name = paste(pathways.show[j], names(object.list)[i], sep = "_")
      title = "Cycling (Top 20 merged pathways)"
    )
  }
}
```



We can plot the gene expression distribution of signaling genes related to L-R pairs or signaling pathway using a Seurat wrapper function `plotGeneExpression`.  
```{r, echo=false}
cellchat@meta$datasets = factor(cellchat@meta$datasets, levels = c("Cycling", "H24", "H48")) # set factor level
plotGeneExpression(cellchat, signaling = "THY1", split.by = "datasets", colors.ggplot = T)
```

```{r}
saveRDS(cellchat, file = "cellchat_timepoint_comparison_2025.rds")
```



