---
title: "Creating_Seurat_Object_v2"
author: "Beth Cragg"
date: "2025-05-26"
output:
  word_document: default
  html_document: default
  pdf_document: default
#Initial data analysis - BC Research Project
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggrepel)
library(viridis)
library(pheatmap)
library(UpSetR)
theme_set(theme_cowplot())
```

Load and process data

```{r, echo=false}
data_dirs <- c(
  "/home/scratch1/msc_2025/s2760053/filtered/2018_control_leukocytes", 
  "/home/scratch1/msc_2025/s2760053/filtered/2018_control_mesenchyme",
  "/home/scratch1/msc_2025/s2760053/filtered/2018_T24_repairing_leukocytes", 
  "/home/scratch1/msc_2025/s2760053/filtered/2018_T24_repairing_mesenchyme",
  "/home/scratch1/msc_2025/s2760053/filtered/2020_control_leukocytes", 
  "/home/scratch1/msc_2025/s2760053/filtered/2020_control_mesenchyme",
  "/home/scratch1/msc_2025/s2760053/filtered/2020_T48_repairing_leukocytes", 
  "/home/scratch1/msc_2025/s2760053/filtered/2020_T48_repairing_mesenchyme"
)

sample_names <- basename(data_dirs)
seurat_objects <- list()

#loop through each dataset
for (i in seq_along(data_dirs)) {
  message("Processing: ", sample_names[i])
  
  data <- ReadMtx(
    mtx = file.path(data_dirs[i], "matrix.mtx.gz"),
    features = file.path(data_dirs[i], "features.tsv.gz"),
    cells = file.path(data_dirs[i], "barcodes.tsv.gz"),
    feature.column = 2
  )
  
  mt.genes <- grep("^mt-", rownames(data), value = TRUE) #mitochdonrial genes
  rp.genes <- grep("^Rp", rownames(data), value = TRUE) #ribosomal genes
  
  percent.mito <- Matrix::colSums(data[mt.genes, , drop = FALSE]) / Matrix::colSums(data)
  percent.ribo <- Matrix::colSums(data[rp.genes, , drop = FALSE]) / Matrix::colSums(data)
  
  #remove mitochondrial and ribosomal genes from the dataset
  data <- data[!rownames(data) %in% mt.genes, ]
  data <- data[!rownames(data) %in% rp.genes, ]
  
  #Create a Seurat object that only includes genes that are expressed in 3 or more cells (with over 200 genes total)
  seu <- CreateSeuratObject(counts = data, min.cells = 3, min.features = 200, project = sample_names[i])
  seu <- AddMetaData(seu, percent.mito, col.name = "percent.mito")
  seu <- AddMetaData(seu, percent.ribo, col.name = "percent.ribo")
  
  seu$timepoint <- if (grepl("T24", sample_names[i])) {
    "24hrs"
  } else if (grepl("T48", sample_names[i])) {
    "48hrs"
  } else {
    "cycling"
  }

  seu$collection <- if (grepl("2018", sample_names[i])) "2018" else "2020"
  seu$celltype <- if (grepl("leukocytes", sample_names[i])) "leukocytes" else "mesenchyme"
  seu$sample <- sample_names[i]
  
  #Scrublet annotation
  doublet_scores <- read.csv(file.path(data_dirs[i], "doublet_scores.csv"))
  predicted_doublets <- read.csv(file.path(data_dirs[i], "predicted_doublets.csv"))
  seu$doublet_score <- doublet_scores$doublet_score
  seu$predicted_doublet <- predicted_doublets$predicted_doublet
  
  # Define a condition to apply a higher threshold for the 2020_control_mesenchyme dataset
  if (sample_names[i] == "2020_control_mesenchyme") {
    seu <- subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito < 0.1 & doublet_score < 0.5)
  } else {
    seu <- subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito < 0.1 & doublet_score < 0.4)
  }

  #Normalisation taken from Phoebe's code - needs checking
  seu <- NormalizeData(seu)
  seu <- FindVariableFeatures(seu, nfeatures = 2000)
  
  #Dimensionality reduction taken from Phoebe's code
  #seu <- ScaleData(seu)
  #seu <- RunPCA(seu, npcs = 20)
  #seu <- RunUMAP(seu, reduction = "pca", dims = 1:20)
  #seu <- FindNeighbors(seu, dims = 1:20)
  #seu <- FindClusters(seu, resolution = 0.2)
  
  seurat_objects[[sample_names[i]]] <- seu
  
  #save individual Seurat objects
  saveRDS(seu, file = paste0("processed_", sample_names[i], ".rds"))
}

```
Integrate Immune and Stromal cell data

```{r, echo=false}
## 🔗 Integration and Clustering by cell type

# Define sample groupings by cell type
leukocyte_samples <- c(
  "2018_control_leukocytes", "2020_control_leukocytes",
  "2018_T24_repairing_leukocytes", "2020_T48_repairing_leukocytes"
)

mesenchyme_samples <- c(
  "2018_control_mesenchyme", "2020_control_mesenchyme",
  "2018_T24_repairing_mesenchyme", "2020_T48_repairing_mesenchyme"
)

# Generalized loader function
load_samples <- function(sample_names) {
  lapply(sample_names, function(name) {
    file_path <- paste0("processed_", name, ".rds")
    message("Reading file: ", file_path)  # This will show the full file path being read
    seu <- readRDS(file_path)
    seu$sample <- name
    return(seu)
  })
}

# Load each group
leukocyte_objects <- load_samples(leukocyte_samples)
mesenchyme_objects <- load_samples(mesenchyme_samples)

# Check if doublet_score is properly added
lapply(leukocyte_objects, function(seu) head(seu@meta.data$doublet_score))
lapply(mesenchyme_objects, function(seu) head(seu@meta.data$doublet_score))


#QC Filtering Function
qc_filter <- function(seu_list) {
  lapply(seu_list, function(seu) {
    subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito < 0.1 & doublet_score < 0.4)
  })
}

#Apply QC Filter
leukocyte_objects <- qc_filter(leukocyte_objects)
mesenchyme_objects <- qc_filter(mesenchyme_objects)


#Integration: Leukocytes
# Find anchors and common genes
anchors_leuk <- FindIntegrationAnchors(object.list = leukocyte_objects, dims = 1:10)
genes_leuk <- Reduce(intersect, lapply(leukocyte_objects, rownames))

# Integrate
leuk_combined <- IntegrateData(anchorset = anchors_leuk, features.to.integrate = genes_leuk)
DefaultAssay(leuk_combined) <- "integrated"

# Dimensionality reduction + clustering
leuk_combined <- ScaleData(leuk_combined)
leuk_combined <- RunPCA(leuk_combined, npcs = 20)
leuk_combined <- RunUMAP(leuk_combined, reduction = "pca", dims = 1:10)
leuk_combined <- FindNeighbors(leuk_combined, dims = 1:10)
leuk_combined <- FindClusters(leuk_combined, resolution = 0.5)

# Save
#saveRDS(leuk_combined, file = "leukocytes_combined.rds")

#Integration: Mesenchyme
anchors_mes <- FindIntegrationAnchors(object.list = mesenchyme_objects, dims = 1:10)
genes_mes <- Reduce(intersect, lapply(mesenchyme_objects, rownames))

mes_combined <- IntegrateData(anchorset = anchors_mes, features.to.integrate = genes_mes)
DefaultAssay(mes_combined) <- "integrated"

mes_combined <- ScaleData(mes_combined)
mes_combined <- RunPCA(mes_combined, npcs = 20)
mes_combined <- RunUMAP(mes_combined, reduction = "pca", dims = 1:10)
mes_combined <- FindNeighbors(mes_combined, dims = 1:10)
mes_combined <- FindClusters(mes_combined, resolution = 0.5)

#saveRDS(mes_combined, file = "mesenchyme_combined.rds")

```
Qaulity control and normalisation checks

```{r, echo = false}
leuk_combined$short_id <- recode(leuk_combined$orig.ident,
  "2018_control_leukocytes" = "18_cont_leuk",
  "2018_T24_repairing_leukocytes" = "24_leuk",
  "2020_T48_repairing_leukocytes" = "48_leuk",
  "2020_control_leukocytes" = "20_cont_leuk")

mes_combined$short_id <- recode(mes_combined$orig.ident,
  "2018_control_mesenchyme" = "18_cont_mes",
  "2018_T24_repairing_mesenchyme" = "24_mes",
  "2020_T48_repairing_mesenchyme" = "48_mes",
  "2020_control_mesenchyme" = "20_cont_mes")

VlnPlot(mes_combined, features = "nFeature_RNA", group.by = "seurat_clusters", pt.size = 0)
VlnPlot(mes_combined, features = "nCount_RNA", group.by = "seurat_clusters", pt.size = 0)
VlnPlot(mes_combined, features = "percent.mito", group.by = "seurat_clusters", pt.size = 0)
VlnPlot(mes_combined, features = "percent.ribo", group.by = "seurat_clusters", pt.size = 0)
VlnPlot(leuk_combined, features = "percent.ribo", group.by = "short_id", pt.size = 0)
# Check if cells with high mito% have low features (potential low-quality)
#FeatureScatter(leukocytes_combined, feature1 = "percent.mt", feature2 = "nFeature_RNA", group.by="short_id")

#checking dimensionality reduction
print(leuk_combined[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(leuk_combined, dims = 1:2, reduction = "pca")
DimPlot(leuk_combined, reduction = "pca") + NoLegend()
DimHeatmap(leuk_combined, dims = 1:5, cells = 500, balanced = TRUE)
ElbowPlot(leuk_combined)  + ggtitle("Immune cell subset")
#Elbow plot shows most SD between 1 and 10

VlnPlot(mes_combined, features = "nFeature_RNA", group.by = "short_id", pt.size = 0)
VlnPlot(mes_combined, features = "nCount_RNA", group.by = "short_id", pt.size = 0)
#VlnPlot(mes_combined, features = "percent.mt", group.by = "short_id", pt.size = 0)
VlnPlot(mes_combined, features = "percent.ribo", group.by = "short_id", pt.size = 0)
# Check if cells with high mito% have low features (potential low-quality)
#FeatureScatter(mes_combined, feature1 = "percent.mt", feature2 = "nFeature_RNA", group.by="short_id")

#checking dimensionality reduction
print(mes_combined[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(mes_combined, dims = 1:2, reduction = "pca")
DimPlot(mes_combined, reduction = "pca") + NoLegend()
DimHeatmap(mes_combined, dims = 1:5, cells = 500, balanced = TRUE)
ElbowPlot(mes_combined)  + ggtitle("Mesenchyme subset")
#Elbow plot shows most SD between 1 and 10

#mes_combined <- RunTSNE(mes_combined, dims = 1:15)
#DimPlot(mes_combined, reduction = "tsne")

```


Check dimensionality reduction and view UMAP

```{r, echo=false}
DimPlot(leuk_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "seurat_clusters") + ggtitle("UMAP of all Leukocytes")
DimPlot(mes_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "seurat_clusters") + ggtitle("UMAP of all Stromal cells")

```
Running a UMAP for all samples

```{r, echo=false}

# Define sample groupings by cell type
all_samples <- c(
  "2018_control_leukocytes", "2020_control_leukocytes",
  "2018_T24_repairing_leukocytes", "2020_T48_repairing_leukocytes",
  "2018_control_mesenchyme", "2020_control_mesenchyme",
  "2018_T24_repairing_mesenchyme", "2020_T48_repairing_mesenchyme"
)


# Generalized loader function
load_samples <- function(sample_names) {
  lapply(sample_names, function(name) {
    seu <- readRDS(paste0("processed_", name, ".rds"))
    seu$sample <- name
    return(seu)
  })
}

# Load each group
all_objects <- load_samples(all_samples)

#QC Filtering Function
qc_filter <- function(seu_list) {
  lapply(seu_list, function(seu) {
    subset(seu, subset = nFeature_RNA > 200 & nFeature_RNA < 4000 & percent.mito < 0.1)
  })
}

#Apply QC Filter
all_objects <- qc_filter(all_objects)

#Integration: All
# Find anchors and common genes
anchors_all <- FindIntegrationAnchors(object.list = all_objects, dims = 1:20)
genes_all <- Reduce(intersect, lapply(all_objects, rownames))

# Integrate
all_combined <- IntegrateData(anchorset = anchors_all, features.to.integrate = genes_all)
DefaultAssay(all_combined) <- "integrated"

# Dimensionality reduction + clustering
all_combined <- ScaleData(all_combined)
all_combined <- RunPCA(all_combined, npcs = 20)
all_combined <- RunUMAP(all_combined, reduction = "pca", dims = 1:20)
all_combined <- FindNeighbors(all_combined, dims = 1:20)
all_combined <- FindClusters(all_combined, resolution = 0.5)

DimPlot(all_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "seurat_clusters") + ggtitle("UMAP of all samples")
FeaturePlot(all_combined, reduction = "umap", features = "Mrc1", ncol=1, pt.size = 0.5, order=TRUE, min.cutoff = 0)

```

Immune Cell Cluster Identification
```{r, echo=false}
#cluster labelling from Phoebe's script

#FeaturePlot(leuk_combined, reduction = "umap", features = cDC2, ncol=4, pt.size = 0.5, order=TRUE, min.cutoff = 0)
leuk_combined$group <- factor(x=leuk_combined$timepoint, levels = c("control", "24hrs", "48hrs"))
FeatureScatter(leuk_combined, feature1 = "nFeature_RNA", feature2 = "Arg1")

#use gene expression to identify clusters
neutrophils <- c("S100a8","S100a9", "Cxcr2" ) #"Lcn2", "Elane", "Mmp8"
macrophages <- c("Itgam", "Zeb2", "Cd68", "Csf1r", "Fcgr1", "Adgre1")
mono <- c("Ly6c2", "Ccr2", "Cd14", "Cx3cr1")
NK <- c("Nkg7", "Gzma", "Gzmb")
DC <- c("Clec9a", "Clec10a", "Cd209a")
Tc <- c("Cd3e", "Cd3d", "Cd3g")
Bc <- c("Cd79a", "Cd79b", "Btla")
#tregs <- c("Foxp3", "Il2ra", "Ctla4")
Eo <- c("Siglecf")


# STEP 1: Identify markers for all clusters
all_markers <- FindAllMarkers(leuk_combined,
                               only.pos = TRUE,
                               min.pct = 0.25,
                               logfc.threshold = 0.25)

# STEP 2: Filter to get top 20 markers per cluster
top_markers <- all_markers %>%
  group_by(cluster) %>%
  slice_max(n = 20, order_by = avg_log2FC)

# STEP 3: Save to CSV for offline review
write.csv(top_markers, "top20_markers_per_cluster.csv", row.names = FALSE)


#FeaturePlot(leuk_combined, reduction = "umap", features = "Siglecf", ncol=1, pt.size = 0.5, order=TRUE, min.cutoff = 0) + ggtitle("Siglecf - Eosinophils")
#FeaturePlot(leuk_combined, reduction = "umap", features = "Ccr3", ncol=1, pt.size = 0.5, order=TRUE, min.cutoff = 0) + ggtitle("Ccr3 - Eosinophils")
#FeaturePlot(leuk_combined, reduction = "umap", features = "Il5ra", ncol=1, pt.size = 0.5, order=TRUE, min.cutoff = 0) + ggtitle("Il5ra - Eosinophils")

#VlnPlot(object = leuk_combined, features = macrophages, ncol = 4, assay = "RNA", pt.size = 0)

DotPlot(leuk_combined, assay = "RNA", features = Eo, cols = c("lavender", "darkblue"), dot.min = 0, col.min = 0, dot.scale = 6, scale.by = "radius") + RotatedAxis()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=8), axis.text.y = element_text((size=10))) + ggtitle("Gene markers for Eosinophils")
```
Stromal Cell Cluster Identification

```{r, echo=false}
#cluster labelling from Phoebe's script

DimPlot(mes_combined, pt.size = 0.5, label.size = 5, label=TRUE, split.by = "timepoint")
mes_combined$timepoint <- factor(mes_combined$timepoint, levels = c("cycling", "24hrs", "48hrs"))


DimPlot(mes_combined, pt.size = 0.5, label.size = 5, label=TRUE)
#VlnPlot(mes_combined, features= c("Pdgfrb", "Pdgfra", "Mcam", "Cspg4", "Hif1a"),  assay="RNA", ncol=5, pt.size=0)
#FeatureScatter(mes_combined, feature1 = "Pdgfrb", feature2 = "Krt18", group.by = "timepoint")

#use gene expression to identify clusters
fibroblasts <- c("Cd34", "Pdgfra", "Mfap4", "Mfap5", "Fbln1", "Fbln2", "Spon2", "Col1a1", "Col3a1")
#"Vim", "Des", "Thy1", "Pdgfrb", "Kcnj8", "Krt8", "Krt19", "Sdc4")
perivascular <- c("Mcam", "Acta2", "Cspg4")
vSMCs <- c("Myh11", "Tagln", "Cnn1", "Kitl", "Pln", "Lmod1")
P1 <- c("Kcnj8", "Abcc9", "Ednrb", "Rgs5", "Vtn", "Postn") # No Myob1
P2 <- c("Arhgdib", "Serpine1", "Cxcl1")
endothelial <- c("Pecam1", "Vwf", "Cdh5", "Cldn5")
epithelial <- c("Cdh1", "Epcam", "Alcam", "Cd24a", "Clu", "Cldn4", "Fxyd3", "Fxyd4", "Krt5", "Krt7", "Krt8", "Krt14", "Krt15", "Krt18", "Krt19", "Muc1", "Muc4", "Wfdc2")
F1 <- c("Aspg", "Dpep1", "Ngfr", "Angptl7", "Ifit1", "Ifit3")
F2 <- c("Cxcl14", "Cdh11", "Rgs2", "Smoc2", "Bmp7", "Wt1")
F3 <- c("Clec3b", "Fap", "Cd55", "Vit", "Col14a1", "Cxcl16")
F4 <- c("Epcam", "Alcam", "Cd24a", "Clu", "Krt18", "Sdc4", "Ndrg1")
F5 <- c("Casp3", "Casp9", "Trp53", "Bax", "Col7a1", "Scube1") #mitochondrial genes and cell death - dead/apoptotic cells which should be excluded

for (i in F4) {
  message("Processing: ", i)
  title <- paste(i,"- Transition")
  fp <- FeaturePlot(mes_combined, reduction = "umap", features = i, ncol=1, pt.size = 0.5, order=TRUE, min.cutoff = 0) + ggtitle(title)
  print(fp)
}

# STEP 1: Identify markers for all clusters
#all_markers <- FindAllMarkers(mes_combined,
 #                              only.pos = TRUE,
  #                             min.pct = 0.25,
   #                            logfc.threshold = 0.25)

# STEP 2: Filter to get top 20 markers per cluster
#top_markers <- all_markers %>%
 # group_by(cluster) %>%
  #slice_max(n = 20, order_by = avg_log2FC)

# STEP 3: Save to CSV for offline review
#write.csv(top_markers, "top20_markers_per_cluster_mes.csv", row.names = FALSE)


#VlnPlot(object = mes_combined, features = macrophages, ncol = 4, assay = "RNA", pt.size = 0)

DotPlot(mes_combined, assay = "RNA", features = F5, cols = c("lavender", "darkblue"), dot.min = 0, col.min = 0, dot.scale = 6, scale.by = "radius") + RotatedAxis()+
  theme(axis.text.x = element_text(angle=45, hjust=1, size=8), axis.text.y = element_text(size=10)) + ggtitle("Gene Markers for F5")

for (tp in unique(mes_combined$timepoint)) {
  subset_obj <- subset(mes_combined, subset = timepoint == tp)
  p <- DimPlot(subset_obj, reduction = "umap", label = TRUE, pt.size = 0.5) +
    ggtitle(paste("UMAP of Stromal Cells at Timepoint:", tp))
  print(p)
}

FeaturePlot(mes_combined, features = c("Cd34", "Mfap5", "Pdgfra"))  # Canonical fibroblasts
VlnPlot(mes_combined, pt.size=0, features = c("Cd34"))  # Canonical fibroblasts
VlnPlot(mes_combined, pt.size=0, features = c("Mfap5"))
VlnPlot(mes_combined, pt.size=0, features = c("Pdgfra"))
mes.markers <- FindAllMarkers(mes_combined, only.pos = TRUE)
mes.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1)
mes.markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 5) %>%
    ungroup() -> top5
DoHeatmap(mes_combined, features = top5$gene) + NoLegend()
```
UMAP of Identified Immune Cells

```{r, echo - false}
#Rename clusters
new_cluster_ids <- c(
  "0" = "Neut",
  "1" = "Mono/Macs",
  "2" = "Neut",
  "3" = "Neut",
  "4" = "NK",
  "5" = "Neut",
  "6" = "Mono/Macs",
  "7" = "Neut",
  "8" = "NK",
  "9" = "Mono/Macs",
  "10" = "DC",
  "11" = "Tc",
  "12" = "Mono/Macs",
  "13" = "DC",
  "14" = "Bc",
  "17" = "Bc"
)

#Apply new identities
leuk_combined <- RenameIdents(leuk_combined, new_cluster_ids)

#Remove clusters 15 and 16
leuk_combined <- subset(leuk_combined, idents = c("15", "16"), invert = TRUE)
leuk_combined$cell_type <- Idents(leuk_combined)
leuk_combined$cell_type_broad <- Idents(leuk_combined)

DimPlot(leuk_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "cell_type") + ggtitle("UMAP of identified Immune Cells")
for (tp in unique(leuk_combined$timepoint)) {
  subset_obj <- subset(leuk_combined, subset = timepoint == tp)
  p <- DimPlot(subset_obj, reduction = "umap", label = TRUE, pt.size = 0.5) +
    ggtitle(paste("UMAP of Immune Cells at Timepoint:", tp))
  print(p)
}
```
UMAP of identified stromal cells

```{r, echo=false}
# View current cluster IDs
Idents(mes_combined) <- "seurat_clusters"
print(levels(Idents(mes_combined)))

#Rename clusters
new_cluster_ids <- c(
  "0" = "F2",
  "1" = "F2",
  "2" = "Peri",
  "3" = "F1",
  "4" = "F1",
  "5" = "vSMCs",
  "6" = "F1",
  "7" = "F2",
  "8" = "F3",
  "9" = "MET",
  "11" = "Epi"
)

mes_combined <- subset(mes_combined, idents = c("10"), invert = TRUE)

# Rename the clusters
Idents(mes_combined) <- "seurat_clusters"
mes_combined <- RenameIdents(mes_combined, new_cluster_ids)
mes_combined$cell_type <- Idents(mes_combined)
table(mes_combined$cell_type, mes_combined$timepoint)


# Define new cluster names
broad_cluster_ids <- c(
  "F2" = "Fibroblasts",
  "F2" = "Fibroblasts",
  "Peri" = "Pericytes",
  "F1" = "Fibroblasts",
  "F1" = "Fibroblasts",
  "vSMCs" = "vSMCs",
  "F1" = "Fibroblasts",
  "F2" = "Fibroblasts",
  "F3" = "Fibroblasts",
  "MET" = "MET",
  "Epi" = "Epithelial"
)

# Rename cluster identities
mes_combined <- RenameIdents(mes_combined, broad_cluster_ids)

# Save these new identities into metadata column
mes_combined$cell_type_broad <- Idents(mes_combined)

# Set broad cell types as the active identity
Idents(mes_combined) <- "cell_type_broad"


DimPlot(mes_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "cell_type_broad") + ggtitle("UMAP of identified Stromal Cells (Broad)")
DimPlot(mes_combined, reduction = "umap", label = TRUE, pt.size = 0.5, group.by = "cell_type") + ggtitle("UMAP of identified Stromal Cells (Specific)")

for (tp in unique(mes_combined$timepoint)) {
  subset_obj <- subset(mes_combined, subset = timepoint == tp)
  p <- DimPlot(subset_obj, reduction = "umap", group.by="cell_type_broad", label = TRUE, pt.size = 0.5) +
    ggtitle(paste("UMAP of Stromal Cells at Timepoint:", tp))
  print(p)
}

```


Visualisation of cell types across timepoints
```{r, echo=false}


# ==== DATA PREPARATION ====

# Standardize and order timepoints for mesenchymal dataset
mes_combined$timepoint[mes_combined$timepoint == "control"] <- "cycling"
mes_combined$timepoint <- factor(mes_combined$timepoint, levels = c("cycling", "24hrs", "48hrs"))

# Standardize and order timepoints for leukocyte dataset
leuk_combined$timepoint <- factor(leuk_combined$timepoint, levels = c("cycling", "24hrs", "48hrs"))


# ==== LEUKOCYTE DATASET ====

# Total cell count per timepoint
leuk_total_counts <- as.data.frame(table(Timepoint = leuk_combined$timepoint))
print("Leukocyte Total Cell Counts per Timepoint:")
print(leuk_total_counts)

# Proportions by cell type and timepoint
leuk_cell_counts <- table(Cell_Type = Idents(leuk_combined), Timepoint = leuk_combined$timepoint)
leuk_cell_proportions <- prop.table(leuk_cell_counts, margin = 2)
df <- table(leuk_combined$cell_type, leuk_combined$short_id)
rowMeans(df[, c("18_cont_leuk", "20_cont_leuk")])



# Format for ggplot
leuk_df <- as.data.frame(as.table(leuk_cell_proportions))
colnames(leuk_df) <- c("Cell_Type", "Timepoint", "Proportion")
leuk_df[is.na(leuk_df$Proportion), "Proportion"] <- 0

# Pie chart
ggplot(leuk_df, aes(x = "", y = Proportion, fill = Cell_Type)) + 
  geom_bar(stat = "identity", width = 1) + 
  coord_polar("y") + 
  facet_wrap(~ Timepoint) + 
  labs(title = "Leukocyte Cell Type Proportions by Timepoint") + 
  theme_void()


# ==== MESENCHYMAL DATASET ====
Idents(mes_combined) <- "cell_type_broad"

# Total cell count per timepoint
mes_total_counts <- as.data.frame(table(Timepoint = mes_combined$timepoint))
print("Mesenchymal Total Cell Counts per Timepoint:")
print(mes_total_counts)

# Proportions by cell type and timepoint
mes_cell_counts <- table(Cell_Type = Idents(mes_combined), Timepoint = mes_combined$timepoint)
mes_cell_proportions <- prop.table(mes_cell_counts, margin = 2)
df <- table(mes_combined$cell_type, mes_combined$short_id)
rowMeans(df[, c("18_cont_mes", "20_cont_mes")])

# Format for ggplot
mes_df <- as.data.frame(as.table(mes_cell_proportions))
colnames(mes_df) <- c("Cell_Type", "Timepoint", "Proportion")
mes_df[is.na(mes_df$Proportion), "Proportion"] <- 0
counts_df <- as.data.frame(as.table(mes_cell_counts))
colnames(counts_df) <- c("Cell_Type", "Timepoint", "Count")
counts_df

# Pie chart
ggplot(mes_df, aes(x = "", y = Proportion, fill = Cell_Type)) + 
  geom_bar(stat = "identity", width = 1) + 
  coord_polar("y") + 
  facet_wrap(~ Timepoint) + 
  theme_void()



```
Merge objects into one

```{r}
merged_seurat <- merge(mes_combined, y=leuk_combined, add.cell.ids = c("mes", "leuk"), merge.data = TRUE)
LayerData(merged_seurat)[1:10,1:15]
merged_seurat[["RNA"]] <-JoinLayers(merged_seurat[["RNA"]])

Idents(merged_seurat) <- "Cell_Type"  # or whatever metadata column holds cell types
subset_cells <- subset(
  merged_seurat, 
  idents = c("Fibroblasts", "MET", "Neut", "Mono/Macs", "Tc", "NK", "Bc", "DC")
)

Idents(subset_cells) <- "timepoint"  # e.g. "Control", "24h", "48h"

control_seurat <- subset(subset_cells, idents = "Control")
h24_seurat     <- subset(subset_cells, idents = "24h")
h48_seurat     <- subset(subset_cells, idents = "48h")

saveRDS(control_seurat, file="2025_MSc_control.rds")
saveRDS(h24_seurat, file="2025_MSc_24.rds")
saveRDS(h48_seurat, file="2025_MSc_48.rds")
```



